Bioconductor workflow for single-cell RNA sequencing: Normalization, dimensionality reduction, clustering, and lineage inference

### 2017-07-06 workflow/workflow.Rmd
Matching cluster colors to previous paper

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
if(packageVersion("zinbwave")<"0.99.6") stop("must have current devel version to avoid bugs")
if(packageVersion("clusterExperiment")<"1.3.2") stop("must have current develop version to avoid bugs")
if(packageVersion("scone")<"1.1.2") stop("must have current devel version to avoid bugs")

library(BiocStyle)
library(knitr)
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=TRUE)
knitr::opts_chunk$set(fig.align="center", fig.width=5, fig.height=5)
```

@@

```{r original}
publishedClusters <- colData(core)[, "publishedClusters"]
col_clus <- c("transparent", brewer.pal(12, "Set3"), brewer.pal(8, "Set2"))
col_clus <- col_clus[1:length(unique(publishedClusters))]
names(col_clus) <- sort(unique(publishedClusters))
table(publishedClusters)
```

@@

```{r mdsW, fig.cap="ZINB-WaVE: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clustering."}
W <- colData(se)[, grepl("^W", colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = "",
     pch = 20, xlab = "Component 1", ylab = "Component 2")
legend(x = "bottomright", legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = "Sample")
```

@@

```{r barplotOurs, fig.cap="RSEC: Barplot of number of cells per cluster for our workflow's RSEC clustering."}
plotBarplot(ceObj, legend = F) 
```

@@

```{r addPublishedClusters, fig.cap="RSEC: Barplot of number of cells per cluster, for our workflow's RSEC clustering, color-coded by original published clustering."}
ceObj <- addClusters(ceObj, colData(ceObj)$publishedClusters, 
                     clusterLabel = "publishedClusters")
plotBarplot(ceObj, whichClusters=c("combineMany","publishedClusters"),
            xlab = "", legend = F) 
```

@@

Cluster name | Description | Color | Correspondence
-------------|-------------|-------| ----------
c1 | HBC | blue | original 1, 5
c2 | GBC | green | original 2, 3, 11
c3 | mSUS | red | original 4, 7
c4 | Contaminants | orange | original -2 
c5 | mOSN | purple | original 9, 12
c6 | Immature Neuron | brown | original 10, 14
c7 | MV | cian | original 15

@@

```{r heatmapsignificant, fig.cap="DE: Heatmap of the normalized expression measures for the 100 most significantly DE genes for the neuronal lineage, where rows correspond to genes and columns to cells ordered by pseudotime."}
topgenes <- names(sort(gam.pval, decreasing = FALSE))[1:100]
heatdata <- y[rownames(se) %in% topgenes, order(t, na.last = NA)]
heatclus <- cl[order(t, na.last = NA)]
ce <- clusterExperiment(heatdata, heatclus, transformation = identity)
plotHeatmap(ce, clusterSamplesData = "orderSamplesValue", breaks = .99)
```

>>>>

```{r options, echo=FALSE, results="hide",message=FALSE, error=FALSE, include=FALSE, autodep=TRUE}
if(packageVersion("zinbwave")<"0.99.6") stop("must have current devel version to avoid bugs")
if(packageVersion("clusterExperiment")<"1.3.2") stop("must have current devel version to avoid bugs")
if(packageVersion("scone")<"1.1.2") stop("must have current devel version to avoid bugs")

library(BiocStyle)
library(knitr)
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE, warning=FALSE)
knitr::opts_chunk$set(fig.align="center", fig.width=5, fig.height=5)
```

@@

```{r original}
publishedClusters <- colData(core)[, "publishedClusters"]
col_clus <- c("transparent", "#1B9E77", "antiquewhite2", "cyan", "#E7298A", "#A6CEE3", "#666666", "#E6AB02", "#FFED6F", "darkorchid2", "#B3DE69", "#FF7F00", "#A6761D", "#1F78B4")
names(col_clus) <- sort(unique(publishedClusters))
table(publishedClusters)
```

@@

```{r mdsW, fig.cap="ZINB-WaVE: MDS of the low-dimensional matrix W, where each point represents a cell and cells are color-coded by original published clustering."}
W <- colData(se)[, grepl("^W", colnames(colData(se)))]
W <- as.matrix(W)
d <- dist(W)
fit <- cmdscale(d, eig = TRUE, k = 2)
plot(fit$points, col = col_clus[as.character(publishedClusters)], main = "",
     pch = 20, xlab = "Component 1", ylab = "Component 2")
legend(x = "topleft", legend = unique(names(col_clus)), cex = .5,
       fill = unique(col_clus), title = "Sample")
```

@@

```{r barplotOurs, fig.cap="RSEC: Barplot of number of cells per cluster for our workflow's RSEC clustering."}
plotBarplot(ceObj, legend = FALSE) 
```

@@

```{r addPublishedClusters, fig.cap="RSEC: Barplot of number of cells per cluster, for our workflow's RSEC clustering, color-coded by original published clustering."}
ceObj <- addClusters(ceObj, colData(ceObj)$publishedClusters, 
                     clusterLabel = "publishedClusters")

## change default color to match with Figure 7
clusterLegend(ceObj)$publishedClusters[, "color"] <- 
  col_clus[clusterLegend(ceObj)$publishedClusters[, "name"]]

plotBarplot(ceObj, whichClusters=c("combineMany","publishedClusters"),
            xlab = "", legend = FALSE)
```

@@

Cluster name | Description | Color | Correspondence
-------------|-------------|-------| ----------
c1 | HBC | blue | original 1, 5
c2 | GBC | green | original 2, 3, 11
c3 | mSUS | red | original 4, 7
c4 | Contaminants | orange | original -2 
c5 | mOSN | purple | original 9, 12
c6 | Immature Neuron | brown | original 10, 14
c7 | MV | cyan | original 15

@@

```{r heatmapsignificant, fig.cap="DE: Heatmap of the normalized expression measures for the 100 most significantly DE genes for the neuronal lineage, where rows correspond to genes and columns to cells ordered by pseudotime."}
topgenes <- names(sort(gam.pval, decreasing = FALSE))[1:100]
heatdata <- y[rownames(se) %in% topgenes, order(t, na.last = NA)]
heatclus <- cl[order(t, na.last = NA)]
ce <- clusterExperiment(heatdata, heatclus, transformation = identity)

#match to existing colors
cols <- clusterLegend(ceObj)$combineMany[, "color"]
names(cols) <- clusterLegend(ceObj)$combineMany[, "name"]
clusterLegend(ce)$cluster1[, "color"] <- cols[clusterLegend(ce)$cluster1[, "name"]]

plotHeatmap(ce, clusterSamplesData = "orderSamplesValue", breaks = .99)
```
